Learning Programs through Bayesian Inference and Planning

### 2013-10-22 drafts/nips/nips2013.tex
Edits to intro.

The vast majority of research in machine learning is focused on utilizing large amounts of noisy data living in high-dimensional, real-valued vector spaces. Such research has been immensely successful, and this has enabled automatic and accurate modelling on a scale that was previously infeasible. However, when we want to build a car, draw a painting, or design the machine learning models to which we apply our new algorithms and techniques, we face problems that current machine learning tools are unable to handle. The highly structured objects that humans produce seems to require a different representation than the ones which machine learning algorithms have most successfully employed.

@@

However, many of the most natural representations of structured domains are best described as programs, are difficult to fit into more restrictive representations.  To take an example directly from the machine learning literature, consider the plate notation~\cite{DBLP:journals/jair/Buntine94} commonly used to describe graphical models over collections of variables. A typical example is shown in Figure~\ref{fig:lda_plate}~\cite{DBLP:journals/jmlr/BleiNJ03}. The goal of plate notation is precisely to express the notion of a ``for loop'' within the graphical model representation. The reason this is a convenient tool is that researchs onceptualize their models as generative programs, not graphs; see the equivalent of the LDA graphical model in the pseudo-code in Figure~\ref{fig:lda_code}. In fact, the field of probabilistic programming, which has gained recent popularity, is spurred on by exactly the motivation that our models are most naturally expressed as programs.

@@

\end{subfigure}%
\begin{subfigure}{.4\textwidth}
	\begin{lstlisting}[frame=single, numbers=none, xleftmargin=0pt]
	for +$k$+ = 1:NumTopics
	  +$\beta$+[k] +$\sim$+ Dirichlet(+$\eta$+)
	for m = 1:NumDocuments
	  +$\theta$+[m] +$\sim$+ Dirichlet(+$\alpha$+) 
	  for n = 1:NumWords 
	    z +$\sim$+ Categorical(+$\theta$+[m])
	   w +$\sim$+ Categorical(+$\beta$+[z])
	\end{lstlisting}
  \caption{The same model as a program.}
  \label{fig:lda_code}
\end{subfigure}
\caption{The popular Latent Dirichlet Allocation model in two forms.}
\label{fig:lda}
\end{figure}

% \begin{figure}[h]
% \begin{center}

% \end{center}
% \caption{A typical example of a graphical model in plate notation. }
% \end{figure}

% \begin{figure}[h]
% \begin{verbatim}

% for k = 1:NumTopics
% 	beta[k]  ~ Dirichlet(eta)
% for m = 1:NumDocuments
% 	theta[m] ~ Dirichlet(alpha) 
% 	for n = 1:NumWords
% 		z ~ Categorical(theta[m])
% 		w[n, m] ~ Categorical(beta[z])

% \end{verbatim}
% \end{figure}

@@

\section{Submission of papers to NIPS 2013}

NIPS requires electronic submissions.  The electronic submission site is  
\begin{center}
   \url{http://papers.nips.cc}
\end{center}

Please read carefully the
instructions below, and follow them faithfully.
\subsection{Style}

Papers to be submitted to NIPS 2013 must be prepared according to the
instructions presented here. Papers may be only up to eight pages long,
including figures. Since 2009 an additional ninth page \textit{containing only
cited references} is allowed. Papers that exceed nine pages will not be
reviewed, or in any other way considered for presentation at the conference.
%This is a strict upper bound.

Please note that this year we have introduced automatic line number generation
into the style file (for \LaTeXe and Word versions). This is to help reviewers
refer to specific lines of the paper when they make their comments. Please do
NOT refer to these line numbers in your paper as they will be removed from the
style file for the final version of accepted papers.

The margins in 2013 are the same as since 2007, which allow for $\approx 15\%$
more words in the paper compared to earlier years. We are also again using 
double-blind reviewing. Both of these require the use of new style files.

Authors are required to use the NIPS \LaTeX{} style files obtainable at the
NIPS website as indicated below. Please make sure you use the current files and
not previous versions. Tweaking the style files may be grounds for rejection.

%% \subsection{Double-blind reviewing}

%% This year we are doing double-blind reviewing: the reviewers will not know 
%% who the authors of the paper are. For submission, the NIPS style file will 
%% automatically anonymize the author list at the beginning of the paper.

%% Please write your paper in such a way to preserve anonymity. Refer to
%% previous work by the author(s) in the third person, rather than first
%% person. Do not provide Web links to supporting material at an identifiable
%% web site.

%%\subsection{Electronic submission}
%%
%% \textbf{THE SUBMISSION DEADLINE IS MAY 31st, 2013. SUBMISSIONS MUST BE LOGGED BY
%% 23:00, MAY 31st, 2013, UNIVERSAL TIME}

%% You must enter your submission in the electronic submission form available at
%% the NIPS website listed above. You will be asked to enter paper title, name of
%% all authors, keyword(s), and data about the contact
%% author (name, full address, telephone, fax, and email). You will need to
%% upload an electronic (postscript or pdf) version of your paper.

%% You can upload more than one version of your paper, until the
%% submission deadline. We strongly recommended uploading your paper in
%% advance of the deadline, so you can avoid last-minute server congestion.
%%
%% Note that your submission is only valid if you get an e-mail
%% confirmation from the server. If you do not get such an e-mail, please
%% try uploading again.

\subsection{Retrieval of style files}

The style files for NIPS and other conference information are available on the World Wide Web at
\begin{center}
   \url{http://www.nips.cc/}
\end{center}
The file \verb+nips2013.pdf+ contains these 
instructions and illustrates the
various formatting requirements your NIPS paper must satisfy. \LaTeX{}
users can choose between two style files:
\verb+nips11submit_09.sty+ (to be used with \LaTeX{} version 2.09) and
\verb+nips11submit_e.sty+ (to be used with \LaTeX{}2e). The file
\verb+nips2013.tex+ may be used as a ``shell'' for writing your paper. All you
have to do is replace the author, title, abstract, and text of the paper with
your own. The file
\verb+nips2013.rtf+ is provided as a shell for MS Word users.

The formatting instructions contained in these style files are summarized in
sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

%% \subsection{Keywords for paper submission}
%% Your NIPS paper can be submitted with any of the following keywords (more than one keyword is possible for each paper):

%% \begin{verbatim}
%% Bioinformatics
%% Biological Vision
%% Brain Imaging and Brain Computer Interfacing
%% Clustering
%% Cognitive Science
%% Control and Reinforcement Learning
%% Dimensionality Reduction and Manifolds
%% Feature Selection
%% Gaussian Processes
%% Graphical Models
%% Hardware Technologies
%% Kernels
%% Learning Theory
%% Machine Vision
%% Margins and Boosting
%% Neural Networks
%% Neuroscience
%% Other Algorithms and Architectures
%% Other Applications
%% Semi-supervised Learning
%% Speech and Signal Processing
%% Text and Language Applications

%% \end{verbatim}

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
preferred typeface throughout. Paragraphs are separated by 1/2~line space,
with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 1/4~inch space above and below title to rules. All pages should
start at 1~inch (6~picas) from the top of the page.

%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.

For the final version, authors' names are
set in boldface, and each name is centered above the corresponding
address. The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

Please pay special attention to the instructions in section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

First level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 12. One line space before the first level
heading and 1/2~line space after the first level heading.

\subsection{Headings: second level}

Second level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the second level
heading and 1/2~line space after the second level heading.

\subsubsection{Headings: third level}

Third level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the third level
heading and 1/2~line space after the third level heading.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone, regardless of the formatter being used.

\subsection{Citations within the text}

Citations within the text should be numbered consecutively. The corresponding
number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
corresponding references are to be listed in the same order at the end of the
paper, in the \textbf{References} section. (Note: the standard
\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
references themselves, any style is acceptable as long as it is used
consistently.

As submission is double blind, refer to your own published work in the 
third person. That is, use ``In the previous work of Jones et al.\ [4]'',
not ``In our previous work [4]''. If you cite your other papers that
are not widely available (e.g.\ a journal paper under review), use
anonymous author names in the citation, e.g.\ an author of the
form ``A.\ Anonymous''.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first footnote} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches
(12~picas).\footnote{Sample of the second footnote}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction; art work should not be
hand-drawn. The figure number and caption always appear after the
figure. Place one line space before the figure caption, and one line
space after the figure. The figure caption is lower case (except for
first word and proper nouns); figures are numbered consecutively.

Make sure the figure caption does not get separated from the figure.
Leave sufficient space to avoid splitting the figure and figure caption.

You may use color figures. 
However, it is best for the
figure captions and the paper body to make sense if the paper is printed
either in black/white or in color.
\begin{figure}[h]
\begin{center}
%\framebox[4.0in]{$\;$}
\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

% \subsection{Tables}

% All tables must be centered, neat, clean and legible. Do not use hand-drawn
% tables. The table number and title always appear before the table. See
% Table~\ref{sample-table}.

% Place one line space before the table title, one line space after the table
% title, and one line space after the table. The table title must be lower case
% (except for first word and proper nouns); tables are numbered consecutively.

% \begin{table}[t]
% \caption{Sample table title}
% \label{sample-table}
% \begin{center}
% \begin{tabular}{ll}
% \multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
% \\ \hline \\
% Dendrite         &Input terminal \\
% Axon             &Output terminal \\
% Soma             &Cell body (contains cell nucleus) \\
% \end{tabular}
% \end{center}
% \end{table}

% \section{Final instructions}
% Do not change any aspects of the formatting parameters in the style files.
% In particular, do not modify the width or length of the rectangle the text
% should fit into, and do not change font sizes (except perhaps in the
% \textbf{References} section; see below). Please note that pages should be
% numbered.

% \section{Preparing PostScript or PDF files}

% Please prepare PostScript or PDF files with paper size ``US Letter'', and
% not, for example, ``A4''. The -t
% letter option on dvips will produce US Letter files.

% Fonts were the main cause of problems in the past years. Your PDF file must
% only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
% to achieve this.

% \begin{itemize}

% \item You can check which fonts a PDF files uses.  In Acrobat Reader,
% select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
% also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
% available out-of-the-box on most Linux machines.

% \item The IEEE has recommendations for generating PDF files whose fonts
% are also acceptable for NIPS. Please see
% \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

% \item LaTeX users:

% \begin{itemize}

% \item Consider directly generating PDF files using \verb+pdflatex+
% (especially if you are a MiKTeX user). 
% PDF figures must be substituted for EPS figures, however.

% \item Otherwise, please generate your PostScript and PDF files with the following commands:
% \begin{verbatim} 
% dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
% ps2pdf mypaper.ps mypaper.pdf
% \end{verbatim}

% Check that the PDF files only contains Type 1 fonts. 
% %For the final version, please send us both the Postscript file and
% %the PDF file.

% \item xfig "patterned" shapes are implemented with 
% bitmap fonts.  Use "solid" shapes instead. 
% \item The \verb+\bbold+ package almost always uses bitmap
% fonts.  You can try the equivalent AMS Fonts with command
% \begin{verbatim}
% \usepackage[psamsfonts]{amssymb}
% \end{verbatim}
%  or use the following workaround for reals, natural and complex: 
% \begin{verbatim}
% \newcommand{\RR}{I\!\!R} %real numbers
% \newcommand{\Nat}{I\!\!N} %natural numbers 
% \newcommand{\CC}{I\!\!\!\!C} %complex numbers
% \end{verbatim}

% \item Sometimes the problematic fonts are used in figures
% included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
% way to clean such figures. For black and white figures, slightly better
% results can be achieved with program \verb+potrace+.
% \end{itemize}
% \item MSWord and Windows users (via PDF file):
% \begin{itemize}
% \item Install the Microsoft Save as PDF Office 2007 Add-in from
% \url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
% \item Select ``Save or Publish to PDF'' from the Office or File menu
% \end{itemize}
% \item MSWord and Mac OS X users (via PDF file):
% \begin{itemize}
% \item From the print menu, click the PDF drop-down box, and select ``Save
% as PDF...''
% \end{itemize}
% \item MSWord and Windows users (via PS file):
% \begin{itemize}
% \item To create a new printer
% on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
% \url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
% AdobePS driver for it to take effect.
% \item To produce the ps file, select ``Print'' from the MS app, choose
% the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
% \item Set ``TrueType Font'' to be ``Download as Softfont''
% \item Open the ``PostScript Options'' folder
% \item Select ``PostScript Output Option'' to be ``Optimize for Portability''
% \item Select ``TrueType Font Download Option'' to be ``Outline''
% \item Select ``Send PostScript Error Handler'' to be ``No''
% \item Click ``OK'' three times, print your file.
% \item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
% the PS file. In Acrobat, check the option ``Embed all fonts'' if
% applicable.
% \end{itemize}

% \end{itemize}
% If your file contains Type 3 fonts or non embedded TrueType fonts, we will
% ask you to fix it.

% \subsection{Margins in LaTeX}

% Most of the margin problems come from figures positioned by hand using
% \verb+\special+ or other commands. We suggest using the command
% \verb+\includegraphics+
% from the graphicx package. Always specify the figure width as a multiple of
% the line width as in the example below using .eps graphics
% \begin{verbatim}
%    \usepackage[dvips]{graphicx} ... 
%    \includegraphics[width=0.8\linewidth]{myfile.eps} 
% \end{verbatim}
% or % Apr 2009 addition
% \begin{verbatim}
%    \usepackage[pdftex]{graphicx} ... 
%    \includegraphics[width=0.8\linewidth]{myfile.pdf} 
% \end{verbatim}
% for .pdf graphics. 
% See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})

% A number of width problems arise when LaTeX cannot properly hyphenate a
% line. Please give LaTeX hyphenation hints using the \verb+\-+ command.

% \subsubsection*{Acknowledgments}

% Use unnumbered third level headings for the acknowledgments. All
% acknowledgments go at the end of the paper. Do not include 
% acknowledgments in the anonymized submission, only in the 
% final paper.

% \subsubsection*{References}

% References follow the acknowledgments. Use unnumbered third level heading for
% the references. Any choice of citation style is acceptable as long as you are
% consistent. It is permissible to reduce the font size to `small' (9-point) 
% when listing the references. {\bf Remember that this year you can use
% a ninth page as long as it contains \emph{only} cited references.}

% \small{
% [1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
% for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
% and T.K. Leen (eds.), {\it Advances in Neural Information Processing
% Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

% [2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
% Realistic Neural Models with the GEneral NEural SImulation System.}
% New York: TELOS/Springer-Verlag.

% [3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
% and recall at excitatory recurrent synapses and cholinergic modulation
% in rat hippocampal region CA3. {\it Journal of Neuroscience}
% {\bf 15}(7):5249-5262.
% }

>>>>

The vast majority of research in machine learning is focused on utilizing large amounts of noisy data living in high-dimensional, real-valued vector spaces. Such research has been immensely successful and has enabled automatic and accurate modelling on a scale that was previously infeasible. However, when we want to build a car, draw a painting, or design the machine learning models to which we apply our new algorithms and techniques, we face problems that current machine learning tools are unable to handle. The highly structured objects that humans produce seems to require a different representation than the ones which machine learning algorithms have most successfully employed.

@@

However, many of the most natural representations of structured domains are best described as programs and are difficult to fit into more restrictive representations.  To take an example directly from the machine learning literature, consider the plate notation~\cite{DBLP:journals/jair/Buntine94} commonly used to describe graphical models over collections of variables. A typical example is shown in Figure~\ref{fig:lda_plate}~\cite{DBLP:journals/jmlr/BleiNJ03}. The goal of plate notation is precisely to express the notion of a ``for loop'' within the graphical model representation. One reason this is a convenient tool is that researchers conceptualize their models as sampling procedures -- easily expressed as programs (see Figure~\ref{fig:lda_code}, but not as static graphs -- and can use plate notation to express a commonly occuring control structure. In fact, the field of probabilistic programming, which has gained recent popularity, is spurred on by exactly the motivation that our models are most naturally expressed as programs.

Suppose, then, that we want to create machine learning algorithms that can help discover good models for data in novel domains; that is, we want algorithms that discovers the structure of the latent graphical model underlying the data. This is a field of active research \

@@

\end{subfigure}%
\begin{subfigure}{.4\textwidth}
	\begin{lstlisting}[frame=single, numbers=none, xleftmargin=0pt]
	for +$k$+ = 1:NumTopics
	  +$\beta$+[k] +$\sim$+ Dirichlet(+$\eta$+)
	for m = 1:NumDocuments
	  +$\theta$+[m] +$\sim$+ Dirichlet(+$\alpha$+) 
	  for n = 1:NumWords 
	    z +$\sim$+ Categorical(+$\theta$+[m])
	    w+$\sim$+ Categorical(+$\beta$+[z])
	\end{lstlisting}
  \caption{The same model as a program.}
  \label{fig:lda_code}
\end{subfigure}
\caption{The popular Latent Dirichlet Allocation model in two forms.}
\label{fig:lda}
\end{figure}

@@
